---
title: "Weight Lifting Exercise Prediction"
author: "Hathairat"
date: "18 May 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

## Executive Summary

This analysis uses machine learning to predict the manner of 6 participants in their exercise. The manner of exercise is represented in 5 classes (A:E). The author created a predicting model using random forest method from the training dataset. The model is later used to predict 20 cases in the test dataset.

## Data Preparation

The training and testing dataset is available in http://groupware.les.inf.puc-rio.br/har. I downloaded the dataset into my working directory and read the data which I called TrainData and TestData.

```{r dataload}
library(dplyr); library(ggplot2); library(caret)
TrainData <- read.csv("~/pml-training.csv")
TestData <- read.csv("~/pml-testing.csv")
```

The Train data has 160 variables and 19,622 observations. I remove any rows that have NAs or blank cells and also the columns with descriptive texts. Eventually the Traindata has 53 columns. Classe which is one of 53 columns is a predictor.

```{r datacleansed}
TrainData[TrainData == ""] <- NA
TrainData <- TrainData[ , colSums(is.na(TrainData)) == 0]
TrainData <- subset(TrainData, select = -c(1:7))
```

A a quick glance and statistical summary of 53 variables which will be used in the analysis are included in *Appendix - Exploratory Data Analysis*.

Before I started prediction models, I split the Train data into testing and training dataset. I also set seed for reproducible purposes. 

```{r datasplit}
inTrain <- createDataPartition(y=TrainData$classe,p=0.7,list=FALSE)
training <- TrainData[inTrain,]
testing <- TrainData[-inTrain,]
set.seed(123)
```

## Creating model with random forests

Since the data has labels, I choose random forest which is one of the supervised predicting methods to build a machine learning model. This is mainly because random forest method can handle classification prediction and returns high accuracy.

```{r MethodRF}
mod1 <- train(classe~., method = "rf", data =training)
pred1 <- predict(mod1, testing)
confusionMatrix(pred1, testing$classe)
```

Here is the model variables ranked by its importance.

``` {r TestImp}
varImp(mod1)
```

### Predicting test data

The model returns high accuracy in the Train data and is expected to give good prediction to the Test dataset. 

Here is the classe prediction of 20 cases.
Note: 1 in the table means predicted classe.

```{r TestRF}
pred1Test <- predict(mod1, TestData)
table(TestData$problem_id,pred1Test)
```

# Appendix

## Exploratory data analysis

Tables below exhibit characteristics and statistic summary of all 53 variables including the prediction - classe from Train dataset.

```{r dataexploratory}
str(TrainData)
```

```{r datastat}
library(psych)
psych::describe(TrainData)
```

